Cryptanalysis of SBC based post-quantum signature scheme

### 2025-05-28 rapport.tex
ajout théorème thibauld feneuil

% !TXS template
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[a4paper]{geometry}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[backend=bibtex]{biblatex}
\usepackage{csquotes}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}
\usepackage{tcolorbox}
\usepackage{enumitem}
\usepackage{fancyvrb}
\usepackage{multicol}
\usepackage{stmaryrd}

@@

		\begin{tcolorbox}[title=Signature Scheme -- Signing algorithm, colback=white, colframe=black]
			\textbf{Inputs:} A public key $x = (\vec{u}, \vec{v})$, a private key $w = (\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ and a message $m \in \{0,1\}^*$.

@@

				\item Compute $X_1, X_2, Y_1, Y_2$ \\
				\quad $X_1 \leftarrow \sum_{i=1}^{N} [X_1^{[e]}]^{\llbracket i \rrbracket}, X_2 \leftarrow \sum_{i=1}^{N} [X_2^{[e]}]^{\llbracket i \rrbracket}, Y_1 \leftarrow \sum_{i=1}^{N} [Y_1^{[e]}]^{\llbracket i \rrbracket}, Y_2 \leftarrow \sum_{i=1}^{N} [Y_2^{[e]}]^{\llbracket i \rrbracket}$

				\item Compute $A^{[e]}$ and $B^{[e]}$ and share them: \\
				\quad $A \leftarrow X_1Y_1 - X_2Y_2$. \\
				\quad $B \leftarrow X_1(\vec{v} \cdot \vec{y}) + Y_1(\vec{u} \cdot \vec{x}) - X_2(\vec{u} \cdot \vec{y}) - Y_2(\vec{v} \cdot \vec{x})$. \\
				\quad $\delta_A \leftarrow A - \sum_{i=1}^{N} [R_A^{[e]}]^{\llbracket i \rrbracket}$. \\
				\quad $\delta_B \leftarrow B - \sum_{i=1}^{N} [R_B^{[e]}]^{\llbracket i \rrbracket}$. \\
				\quad $\vec{\delta}_x \leftarrow A - \sum_{i=1}^{N} [\vec{R}_x^{[e]}]^{\llbracket i \rrbracket}$. \\
				\quad $\vec{\delta}_y \leftarrow A - \sum_{i=1}^{N} [\vec{R}_y^{[e]}]^{\llbracket i \rrbracket}$.

@@

>>>>

% !TXS template
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[a4paper]{geometry}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[backend=bibtex]{biblatex}
\usepackage{csquotes}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}
\usepackage[most]{tcolorbox}
\usepackage{enumitem}
\usepackage{fancyvrb}
\usepackage{multicol}
\usepackage{stmaryrd}

@@

		\begin{tcolorbox}[title=Protocol 15: Signing Algorithm, breakable, enhanced, colback=white, colframe=black]
			\textbf{Inputs:} A public key $x = (\vec{u}, \vec{v})$, a private key $w = (\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ and a message $m \in \{0,1\}^*$.

@@

				\item Compute $X_1^{[e]}, X_2^{[e]}, Y_1^{[e]}, Y_2^{[e]}$ \\
				\quad $X_1^{[e]} \leftarrow \sum_{i=1}^{N} [X_1^{[e]}]^{\llbracket i \rrbracket}, X_2^{[e]} \leftarrow \sum_{i=1}^{N} [X_2^{[e]}]^{\llbracket i \rrbracket}, Y_1^{[e]} \leftarrow \sum_{i=1}^{N} [Y_1^{[e]}]^{\llbracket i \rrbracket}, Y_2^{[e]} \leftarrow \sum_{i=1}^{N} [Y_2^{[e]}]^{\llbracket i \rrbracket}$

				\item Compute $A^{[e]}$ and $B^{[e]}$ and share them: \\
				\quad $A^{[e]} \leftarrow X_1Y_1 - X_2Y_2$. \\
				\quad $B^{[e]} \leftarrow X_1(\vec{v} \cdot \vec{y}) + Y_1(\vec{u} \cdot \vec{x}) - X_2(\vec{u} \cdot \vec{y}) - Y_2(\vec{v} \cdot \vec{x})$. \\
				\quad $\delta_A^{[e]} \leftarrow A^{[e]} - \sum_{i=1}^{N} [R_A^{[e]}]^{\llbracket i \rrbracket}$. \\
				\quad $\delta_B^{[e]} \leftarrow B^{[e]} - \sum_{i=1}^{N} [R_B^{[e]}]^{\llbracket i \rrbracket}$. \\
				\quad $\vec{\delta}_x^{[e]} \leftarrow A^{[e]} - \sum_{i=1}^{N} [\vec{R}_x^{[e]}]^{\llbracket i \rrbracket}$. \\
				\quad $\vec{\delta}_y^{[e]} \leftarrow A^{[e]} - \sum_{i=1}^{N} [\vec{R}_y^{[e]}]^{\llbracket i \rrbracket}$.

				\item Compute random value $t_0^{[e]}$ and share it. \\
				\quad $[t_0^{[e]}]^{\llbracket i \rrbracket} \leftarrow Hash_0(i || [vec{R}_x^{[e]}]^{\llbracket i \rrbracket}, [\vec{R}_y^{[e]}]^{\llbracket i \rrbracket}, [X_1^{[e]}]^{\llbracket i \rrbracket}, [X_2^{[e]}]^{\llbracket i \rrbracket}, [Y_1^{[e]}]^{\llbracket i \rrbracket}, [Y_2^{[e]}]^{\llbracket i \rrbracket},$ \\ $[R_A^{[e]}]^{\llbracket i \rrbracket}, [R_B^{[e]}]^{\llbracket i \rrbracket} || \vec{\delta}_x^{[e]}, \vec{\delta}_y^{[e]}, \delta_A^{[e]}, \delta_B^{[e]})$. \\
				\quad $t_0^{[e]} \leftarrow \sum_{i=1}^{N} [t_0^{[e]}]^{\llbracket i \rrbracket}$.

@@

	\begin{theorem}[\cite{feneu2023}]
		Let $\text{Hash}_0, \text{Hash}_1, \text{Hash}_m$ and $\text{Expand}$ be modeled as random oracles, and let $(N, \tau, \lambda, p)$ be parameters of the signature scheme. Let $\mathcal{A}$ be an adversary against the \textbf{EUF-CMA} security of the scheme running in time $t_{\mathcal{A}}$ and making a total of $Q_{RO}$ random oracle queries and $Q_{Sig}$ signing queries. Assuming that $F$ is an $(t_{\mathcal{A}}, \epsilon_{OWF})$-hard one-way function and that PRG is a $(t_{\mathcal{A}}, \epsilon_{PRG})$-secure psuedorandom generator then $\mathcal{A}$'s advantage in the \textbf{EUF-CMA} game is
		$$
			\epsilon_{EUF-CMA} \leq \epsilon_{OWF} + \epsilon_{PRG} + \frac{(\tau N + 2)Q^2}{2^{2\lambda}} + Pr[X+Y = \tau],
		$$
		where $Q = Q_{RO} + N_{Hash} \cdot Q_{Sig}$ with $N_{Hash} = 2 + \tau(2N - 1)$ the number of hash computations in a signature generation, and where $X, Y$ are defined as
		\begin{itemize}
			\item $X = max_{q1 \in \mathcal{Q}_1} \{X_{q1}\} \text{ with } X_{q1} \sim \mathcal{B}\left(\tau, \binom{N}{l+1}p\right)$ and
			\item $Y = max_{q2 \in \mathcal{Q}_2} \{Y_{q2}\} \text{ with } Y_{q2} \sim \mathcal{B}\left(\tau - X, 1 / \binom{N}{l}\right)$,
		\end{itemize}
		where $\mathcal{B}(n_0, p_0)$ denotes the binomial distribution with $n_0$ the number of trials and $p_0$ the success probability of each trial.
	\end{theorem}
